Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores with Time Series Modeling

### 2020-02-07 Ben_Manuscripts/stochastic_transport/cover_letter.tex
HDP-HMM outline

	Recently, we characterized transport mechanisms exhibited by a series of 
	small polar solutes in an LLC Membrane (Coscia and Shirts \textit{J. Phys.
	Chem. B}, 123, 6314--6330 (2019)). We observed subdiffusive solute transport
	characterized by intermittent hops separated by periods of entrapment. In
	this work, we build on the chemical intuition gained from our previous study
	in order to formulate stochastic models which produce trajectory realizations
	that share the same dynamical properties as those exhibited by solutes in 
	our MD simulations. We explore two modeling approaches. First, based on the
	anomalous diffusion literature, we hypothesize that solutes are subject 
	to subordinated fractional Brownian or L\'evy motion. This implies that 
	solutes experience a series of dwell times separated by anti-correlated hop
	lengths. Our second approach adds state-dependent dynamics to standard Markov
	state modeling approaches. We model solute motion as a series of transitions
	between various trapped states, each with different dynamical properties.

%        Recently, we characterized an atomistic molecular model of an
%        LLC membrane that is maximally consistent with experimental
%        observations (Coscia et al. \textit{J. Phys. Chem. B}, 123,
%        289--309 (2019)). In this work, we observe and quantify
%        transport of water, sodium and 20 small polar solutes within
%        the pores of our atomistic H\textsubscript{II} phase LLC
%        membrane model. In general, all solutes exhibit subdiffusive
%        transport behavior characterized by intermittent hops between
%        long periods of entrapment. However, due the pores'
%        inhomogeneous architecture, we observed 3 different mechanisms
%        of entrapment. First, solutes can diffuse out of the pores and
%        become entangled between tails. Second, many of the solutes
%        are capable of donating hydrogen bonds to monomer head groups
%        for extended periods of time. Finally, solutes can associate
%        with monomer head group counter-ions within nanopores. Each
%        solute is influenced by each mechanism to varying degrees
%        dependent on its size and chemical functionality. We begin our
%        discussion with a broad description of these mechanisms in the
%        context of all solutes before restricting the discussion to
%        more detailed analyses of subsets of chemically-similar
%        solutes.

>>>>

	Recently, we characterized transport mechanisms exhibited by a series of 
	small polar solutes in an LLC Membrane (Coscia and Shirts \textit{J. Phys.
	Chem. B}, 123, 6314--6330 (2019)). We observed subdiffusive solute transport
	characterized by intermittent hops separated by periods of entrapment. In
	this work, we build on the chemical intuition gained from our previous study
	in order to formulate stochastic models which produce trajectory realizations
	that share the same dynamical properties as those exhibited by solutes in 
	our MD simulations. We explore two modeling approaches. First, based on the
	anomalous diffusion literature, we parameterize solute motion as 
    subordinated fractional Brownian and L\'evy motion processes. This implies that 
	solutes experience a series of dwell times separated by anti-correlated hop
	lengths. Our second approach adds state-dependent dynamics to standard Markov
	state modeling approaches. We model solute motion as a series of transitions
	between various trapped states, each with different dynamical properties.
	Finally, we demonstrate how one can project our best model onto macroscopic 
	timescales in order to estimate solute flux and selectivity.
